\speculake: Hiding Malware in Speculative Execution

### 2018-06-21 paper/appendix_spasm.tex
VM implementation updates
target fn translated to speculative gadget

The progamming model used by SPASM builds up values on the maintained stack
and stores values in specially reserved offsets equivalent to x86 registers 
in order to perform system calls. Immediates can be assigned into VAL
incrementally, and stored, operated on, or pushed onto the maintained
stack. Control Flow, including conditional jmps, is implemented by modifying 
the maintained instruction pointer SRIP.

Using this we have implemented various example programs including 
\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{DialbackShell}.
See Figure~\ref{fig:spasm_model} for a diagram of the SPASM emulator's
usage within the \speculake model.

>>>>

The progamming model used by SPASM builds up values on the maintained stack
and stores them in specially reserved offsets equivalent to x86 registers 
in order to perform system calls. Using the \texttt{UVAL} instruction
an immediate can be assigned into the lowest 4 bits of VAL, which 
can subsequenty be shifted to allow futher bits to be entered. In this way
immediates can be built up incrementally, stored, operated on, or pushed 
onto the maintained stack. Control Flow, including conditional jumps, 
is implemented by modifying the maintained instruction pointer SRIP.

Using this we have implemented various example programs including 
\textit{HelloWorld}, \textit{FizzBuzz}, and \textit{ReverseShell}.
See Figure~\ref{fig:spasm_model} for a diagram of the SPASM emulator's
usage in an \speculake application.

### 2018-06-21 paper/architecture.tex

>>>>

\subsubsection{Nested Speculation}

We explore the ability for the CPU to ``double speculate'', where a second
stalled indirect jump while the CPU is already speculating causes it to predict
the target and
speculate a second time. For instance, suppose a payload program truly jumps to
target A, but the CPU is mistrained by a trigger program that jumps to B, thus
causing the payload program to speculatively execute at B. At B, suppose there
is a second indirect jump, perhaps using the same register as the first jump
(which has still not resolved). If the trigger program jumps to C, the payload
program will speculate a second time and continue speculative execution at C.

We find that not all Intel CPUs support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation. We find that when a CPU does support nested speculation, there
appears to be no limit to nested depth besides the speculative instruction
limit. We use a 16-deep nested speculation in Section~\ref{subsec:impl-aes}
to thoroughly protect speculative decryption keys from reverse
engineering.

### 2018-06-21 paper/background.tex
partial related work, and bib fixes across paper

In early 2018, researchers revealed the Spectre vulnerability, which allows an
attacker to leak information from a victim program~\cite{specte}. Spectre uses
the fact that speculative execution can influence system state via side
channel. In Spectre, an attacker mistrains the branch predictor of a CPU running
a victim program by providing inputs to it. Once mistrained, the attacker then
sends a new input that will cause a different in-order execution path. However,
because the CPU's branch predictor has been mistrained, it will still
speculatively execute the previous path.

>>>>

In early 2018, researchers revealed the Spectre vulnerability, which allows an
attacker to leak information from a victim
program~\cite{spectre,project_zero,maisuradze2018speculose}. Spectre uses the
fact that speculative execution can influence system state via side channel. In
Spectre, an attacker mistrains the branch predictor of a CPU running a victim
program by providing inputs to it. Once mistrained, the attacker then sends a
new input that will cause a different in-order execution path. However, because
the CPU's branch predictor has been mistrained, it will still speculatively
execute the previous path.

### 2018-06-21 paper/conclusion.tex

In this paper we have presented \speculake, a model for hiding computation in
speculative execution that is fundamentally different than modern methods of
obfuscating computation. Through a series of experiments we have classified the
capabilities and limitations of this speculative primitive and demonstrated
various example applications. We have demonstrated the widely varying
capabilities of speculative execution, in particular showing that it allows for
arbitrary computation from a Turing Machine, emulation, transparent and remote
communication, and process level data encryption. As a result current analysis
techniques for detecting process cooperation, and describing the behavior of
program binaries are no longer definitively reliable.

@@

Software isolation techniques are critical in protecting against exploitation
models, however, complex optimization has often neglected these principles in
favor of performance. The stop-gap measures proposed in this work contain
security and performance trade-offs, though we advocate for a more robust
solution in future processor generations. Long term solutions may require that
we take a step backward before we can securely step forward.

>>>>

%In this paper 
We have presented \speculake, a model for hiding computation in
speculative execution that is fundamentally different than existing methods of
code obfuscation. Through a series of experiments we have classified the
capabilities and limitations of this speculative primitive and demonstrated
various example applications. We have demonstrated the potential
of using speculative execution in several applications, including a Turing
machine, SPASM emulator, remotely-triggered payloads, and AES decryption.
Currently analysis techniques for reverse engineering are insufficient to reason
about the behavior of these programs.

@@

Ultimately, silicon and microarchitecture patches will be needed to secure CPUs
against this kind of malware. Meanwhile, attackers may
iterate and find new variants of \speculake-like malware.
In the meantime, new detection techniques and
software-level mitigations are desperately needed.

%Software isolation techniques are critical in protecting against exploitation
%models, however, complex optimization has often neglected these principles in
%favor of performance. The stop-gap measures proposed in this work contain
%security and performance trade-offs, though we advocate for a more robust
%solution in future processor generations. Long term solutions may require that
%we take a step backward before we can securely step forward.
%

### 2018-06-21 paper/figures.tex

    \caption{The number of instructions and error rate of a the 5-state Busy
            Beaver configuration were measured as a function of how many
            (redundant) iterations into the speculative world were performed. As
            expected, the amount of time to complete a million instructions is a
            linear function of how many indirect calls into the speculative
            world we take. Additionally we see that it only requires using 10
            speculative calls per instruction to achieve a very low number of
            errors in reaching a million instructions, making 10-20 iterations
            an attractive choice for most applications.}

>>>>

    \caption{The number of steps and error rate of a the 5-state Busy
            Beaver configuration were measured as a function of how many
            (redundant) iterations into the speculative world were performed. As
            expected, the amount of time to complete a million steps is a
            linear function of how many indirect calls into the speculative
            world we take. Additionally we see that it only requires using 10
            speculative calls per step to achieve a very low number of
            errors in reaching a million instructions, making 10-20 steps
            an attractive choice for most applications.}

### 2018-06-21 paper/implementation.tex

In this model of computation there are no instruction arguments, as we must
return an entire instruction from the speculative world inside the limited-width
cache side channel. Although other small
instruction sets exist, they either allow variable instruction
lengths, are too long even in reduced form, or did not have significant support
to make them favorable for developers.

We have used 6 bits in the construction of this instruction set as our
goal is to limit the length of each opcode as much as possible. Note that this
is different from the goal in maximizing bandwidth, as our goal now is to
maximize instruction throughput.

In SPASM we have implemented multiple example programs that can be run as
encrypted binaries in an \speculake payload: First, a \textit{HelloWorld}
program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
demonstrates control flow and arithmetic operations while printing to stdout.
And finally, a \textit{ReverseShell} program that opens and connects to a socket
before  duplicating I/O file descriptors and executing a local shell. Details of
the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}. Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

Expected performance for a given SPASM binary will vary given multiple factors,
the most important of which is the \texttt{Flush+Reload} redundancy. This can be
tuned based on the specific trigger, as it is used to establish confidence that
a signal has been identified in the indices returned from the "speculative
world". 
%Thus accomplishing a task using SPASM generally equates to:

@@

For example, the \textit{ReverseShell} program is 355 SPASM instructions, and
makes six system calls to open a socket,
connect to it, duplicate I/O file descriptors, and perform an \texttt{execve}
system call to open a shell.
In our tests using 5 iterations per decrypted instruction, the
\textit{ReverseShell} program takes just over 2ms to launch a reverse shell
once triggered.

@@

Benign programs can act as a trigger program, provided they perform sufficient
indirect jumps to train the branch predictor. We experiment using the OpenSSL
library as a potential benign trigger, as its source code has gratuitous use of
function pointers which compile to indirect jumps. In addition, it has many
complicated code paths that can be easily selected by remote clients by their
choice of cipher suite.

>>>>

In this model of computation there are effectively no instruction arguments, as we must
return an entire instruction from the speculative world inside the limited-width
cache side channel. Although other small
instruction sets exist, they either allow variable instruction
lengths, are too long even in reduced form, or did not have significant support
to make them favorable for developers.

We used 6 bits in the construction of this instruction set as our
goal is to limit the length of each opcode as much as possible. Note that this
is different from the goal in maximizing bandwidth, as our goal now is to
maximize instruction throughput. Given our short instructions, loading values into registers
requires shifting in 4-bits at a time. SPASM has two registers that act as a
pointer and working register, that can be used to perform jumps, arbitrary
memory reads and writes, and basic arithmatic. We also have a \texttt{syscall}
instruction that makes a real system call to the underlying operating system
with parameters loaded from the SPASM state, allowing us to interact with the
real world. Details of
the SPASM instruction set can be found in Appendix~\ref{appendix:spasm}.

In SPASM we have implemented multiple example programs that we encrypted and
loaded into a \speculake payload, which decrypts and emulates SPASM instructions
only when the corresponding trigger program is running.
We first implemented a \textit{HelloWorld}
program that prints to \texttt{stdout}. Next, a \texttt{FizzBuzz} program that
demonstrates control flow and arithmetic operations while printing to stdout.
Finally, we implemented a \textit{ReverseShell} program that opens and connects to a socket
before duplicating I/O file descriptors and executing a local shell.
Figure~\ref{fig:spasm_model} details the high-level flow of a SPASM payload.

%Expected performance for a given SPASM binary will vary given multiple factors,
%the most important of which is the \texttt{Flush+Reload} overhead. This can be
%tuned based on the specific trigger, as it is used to establish confidence that
%a signal has been identified in the indices returned from the "speculative
%world". 
%Thus accomplishing a task using SPASM generally equates to:

@@

Our \textit{ReverseShell} program consists of 355 SPASM instructions, and
makes six system calls to open a socket,
connect to it, duplicate I/O file descriptors, and perform an \texttt{execve}
system call to open a shell.
In our tests using 5 iterations per decrypted instruction, the
\textit{ReverseShell} program takes just over 2ms to launch a reverse shell
once triggered.

@@

Benign programs can also act as a trigger program, provided they perform sufficient
indirect jumps to train the branch predictor. We experiment using the OpenSSL
library as a potential benign trigger, as its source code has gratuitous use of
function pointers which compile to indirect jumps. In addition, it has many
complicated code paths that can be easily selected by remote clients by their
choice of cipher suite.

### 2018-06-21 paper/introduction.tex

Using these building blocks, we demonstrate the practicality of hiding arbitrary
computation by implementing a reverse shell in our speculative instruction set.
We show that this simple payload is able to perform several system calls
in a reasonable time, ultimately launching a dial-back TCP shell in under 7~seconds after
the trigger program starts.

>>>>

Using these building blocks, we demonstrate the practicality of hiding arbitrary
computation by implementing a reverse shell in our speculative instruction set.
We show that this simple payload is able to perform several system calls in a
reasonable time, ultimately launching a dial-back TCP shell in just over
2~milliseconds after the trigger program starts.

### 2018-06-21 paper/payloads.tex
implementation pass

The computation of the Turing Machine is done speculatively, meaning that while
the real world keeps track of the state, tape, and symbols of the turing machine
all logic used to update the above is done in the speculative world.

@@

\subsubsection{Nested Speculation}

We explore the ability for the CPU to ``double speculate'', where a stalled
indirect jump while the CPU is speculating causes it to predict the target and
speculate a second time. For instance, suppose a payload program truly jumps to
target A, but the CPU is mistrained by a trigger program that jumps to B, thus
causing the payload program to speculatively execute at B. At B, suppose there
is a second indirect jump, perhaps using the same register as the first jump
(which has still not resolved). If the trigger program jumps to C, the payload
program will speculate a second time and also begin executing code at C.

We find that not all Intel CPUs support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation.
% Could cut
We leverage this support in Section~\ref{subsec:decryption} to
increase the security of speculative decryption keys.

%There are numerous challenges for a reverse engineer to overcome to learn how a
%malicious program with this primitive works. As with previous examples, the
%malicious instructions are only executed when the correct trigger program is
%running (priming the jump table predictor \textbf{TK: word for this?}). However,
%when the to-be-executed instructions are executed the reverse engineer cannot
%locate these instructions in the ELF file, or generated source code. These
%instructions (when they even appear) seem to be generated by themselves, with no
%apparent cause. Additionally, there is no indicator that these instructions are
%encrypted as the decryption is done in the speculative world, meaning that no
%decryption (or key interaction) is done in a way visible to a reverse engineer.
%
%This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).

@@

We design a custom emulator and instruction set---SPASM (Speculative
Assembly)---that accommodates the constraints of the speculative primitive.
SPASM is a 6-bit Instruction set, where all instructions (including operand,
registers, and arguments) fit within 6-bits. This allows each step of the
speculative world to emit a single SPASM instruction to the real world for
emulation by a light-weight SPASM emulator. Using SPASM, developers can write
programs, and assemble and encrypt them into a payload program. When the
associated trigger program runs, the payload will decrypt SPASM instructions in
the speculative world, and execute them one at a time.

>>>>

The computation of the Turing Machine is done speculatively, meaning that while
the real world keeps track of the state, tape, and symbols of the turing machine
all logic used to update the above is done in the speculative world. Figure~\ref{fig:general_model} demonstrates the execution flow of a sample \speculake malware.

@@

% Could cut
%We leverage this support in Section~\ref{subsec:decryption} to
%increase the security of speculative decryption keys.
%
%There are numerous challenges for a reverse engineer to overcome to learn how a
%malicious program with this primitive works. As with previous examples, the
%malicious instructions are only executed when the correct trigger program is
%running (priming the jump table predictor \textbf{TK: word for this?}). However,
%when the to-be-executed instructions are executed the reverse engineer cannot
%locate these instructions in the ELF file, or generated source code. These
%instructions (when they even appear) seem to be generated by themselves, with no
%apparent cause. Additionally, there is no indicator that these instructions are
%encrypted as the decryption is done in the speculative world, meaning that no
%decryption (or key interaction) is done in a way visible to a reverse engineer.
%
%This method allows for arbitrary computation that is unobservable to any observer (static or dynamic).

@@

We design a custom emulator and instruction set---SPASM (Speculative
Assembly)---that accommodates the constraints of the speculative primitive.
SPASM is a 6-bit Instruction set, where all instructions (including operand,
registers, and arguments) fit within 6-bits. This allows each step of the
speculative world to emit a single SPASM instruction to the real world for
emulation by a light-weight SPASM emulator. Using SPASM, developers can write
programs, assemble and encrypt them into a payload program. When the
associated trigger program runs, the payload will decrypt SPASM instructions in
the speculative world, and execute them one at a time.
